Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys

### 2019-01-24 doc/article.tex
forging on

In the next decade, SNe~Ia will be used as peculiar-velocity probes to measure  the influence of gravity on structure formation within the Universe.
The peculiar velocity power spectrum is sensitive to the growth of structure as $P_{vv}\propto (fD)^2$, where $D$ is  the linear growth factor and
$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.\footnote{
To be precise, the peculiar
velocity power spectrum has dependence $P_{vv}\propto (HfD)^2$.  A supernova  survey  measures luminosity distance fluctuations
$\delta_{d_L} = (d_L- \bar{d}_L(z))/\bar{d}_L(z)$, where $d_L$ is the observed distance and $\bar{d}_L(z)$ is the expected distance at the observed redshift $z$.
To first order in peculiar velocity along the line of sight $v$, $\delta_{d_L} =  v \left(1-\frac{1}{H\bar{d}_L(z)}\right) \approx -\frac{v}{H\bar{d}_L(z)}$
at low redshift.   The $H$-dependence of  $P_{vv}$ and the conversion from distances to velocities cancel, making  peculiar velocity surveys sensitive to $(fD)^2$.
}
The $\Lambda$CDM prediction for the $z=0$ velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
\citet{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravities 
follow the relation
$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively.  Adopting this parameterization to model gravity, peculiar velocity
surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose $\Lambda$CDM prediction is plotted 
in Figure~2 of  \citet{1475-7516-2013-04-031}.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{src/new2.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{Volume-weighted velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu=1, 0.5$ (solid black, cyan)  as predicted for General Relativity in the linear regime.
Volume-weighted peculiar velocity shot-noise  $k^3\sigma^2/n$ at $z=0.1$, expected from 2- and 10-year (red, black) supernova densities, 0.08 and 0.15~mag (dashed, dotted) intrinsic magnitude dispersions.
The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
redshift depths $z_{max}$.
\label{power:fig}}
\end{figure}

@@

Two advances in the upcoming decade will make SN~Ia peculiar velocities more powerful.
First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\sigma_M \gtrsim 0.12$~mag absolute magnitude
dispersion.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
Recent studies indicate that with the right data, SNe absolute
magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.  Though not yet
established, it is anticipated that such a reduction in intrinsic dispersion comes with a reduction in the magnitude bias correlated with host-galaxy properties
that is observed using current calibrations.
One such SN is worth $\gtrsim 25$ galaxies with 0.4~mag absolute magnitude uncertainty.
Secondly,  in the upcoming decade cadenced wide-field imaging surveys such as ZTF and LSST
  will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia
 for which good light curves can be measured, corresponding to a  number density of $n \sim 5\times 10^{-4}h^3$\,Mpc$^{-3}$
 (the total rest-frame volumetric SN~Ia rate being $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$.)
  This sample has comparable
 number density and more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.

@@

Our scientific interest is in probing gravity, so while most of the literature cast peculiar velocity results in terms of 
$fD$, we now focus on $\gamma$.  Survey design should be tailored toward minimizing the uncertainty in $\gamma$
$\sigma_\gamma$ rather than $\sigma_{\ln{fD}}$.  To illustrate the distinction,
 $\frac{d(\ln{fD})}{d\gamma} = \ln\Omega + \int \Omega^{\gamma} \ln{\Omega} d\ln{a}$, which for $\Lambda$CDM
 $\approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ respectively; given two surveys with the same percentage precision  in $fD$, the one at lower redshift
better constrains $\gamma$. 
The uncertainties in $\gamma$ are estimated using a
Fisher matrix analysis
similar  to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}.
The analysis is performed in Fourier space, 
where the Fisher information matrix of parameters $\lambda \in \{\gamma,bD\}$ is
\begin{align}
F_{ij} 
%& = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
%\frac{\partial C}{\partial \lambda_j} \right]\\
& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
\label{fisher:eqn}
\end{align}
where
\begin{equation}
C(k,\mu)  =
  \begin{bmatrix}
   P_{\delta \delta}(k,\mu) + \frac{1}{n} &
   P_{v\delta}(k,\mu)  \\
   P_{v\delta}(k,\mu)  &
  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
   \end{bmatrix}.
\label{cov:eq}
\end{equation}
The parameter dependence enters through $fD$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia overdensity
power spectrum $P_{\delta \delta }\propto (bD + fD\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
\propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu$ is the cosine of the angle between the $k$-mode
and the line of sight.  Note that while the $bD$ term does contain information on $\gamma$, we do not use it in this analysis.
The uncertainty in $\gamma$ is $\sigma_\gamma = \sqrt{\left(F^{-1}\right)_{\gamma,\gamma}}$.
Non-GR models may  also predict a change in the scale-dependence of the growth or non-constant $\gamma$, such observations provide additional leverage in probing gravity
but are not considered here.

The survey parameters that enter the calculation are the solid angle $\Omega$, depth of the survey $r_{max}$ (or $z_{max}$), the intrinsic SN~Ia magnitude dispersion 
through $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$, and the survey duration $t$ through $n=\epsilon rt$ where $r$ is the observer SN~Ia rate
and $\epsilon$ is the sample-selection efficiency.
An analysis in which the RSD and velocity surveys are treated independently is considered by
setting the off-diagonal elements of $C$ to zero.

We calculate  $\sigma_\gamma$ of  peculiar velocity surveys for a range of redshift depths $z_{max}$, with survey durations of $t=2$ and 10~years, covering $3\pi$  of sky, and with 
analyses accounting for or neglecting
cross-correlations.
The other survey parameters
$\Omega=3$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
The bias of SN~Ia hosts is taken to be $b=1.2$.
The uncertainties in $\gamma$ are shown in Figure~\ref{var:fig}.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{src/var.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{Projected $\sigma_\gamma$ of  peculiar velocity surveys for a range of redshift depths $z_{max}$, with survey durations of $t=2$ and 10~years, and with 
analyses accounting for or neglecting
cross-correlations.
\label{var:fig}}
\end{figure}

Shallow, short-duration peculiar-velocity surveys already achieve meaningful constraints on gravity.   A 2-year, $z_{max}=0.1$ survey has $\sigma_\gamma \sim 0.038$,
which can distinguish  between  General Relativity, $f(R)$,  and DGP gravities at the $2\sigma$ level.
Using cross-correlation information decreases $\sigma_\gamma$ by  factors of $\sim 0.8$.   
 This modest improvement 
comes from the suppression of  sample variance.
It is the density measurement that is sample-variance
limited,  the peculiar velocity measurement is shot-noise limited.

Having established that peculiar velocity surveys can produce a precise test of gravity, we now consider survey optimization through the 
survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also show the relative importance of different
redshifts with a given survey.  An important survey parameter that is not considered is cost.
\begin{description}
\item[Solid Angle $\Omega$]
The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$.  The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole could significantly increase
the sky coverage with a corresponding decrease in the variance of $\gamma$.
\item[Differential Redshift Bin $z$] Certain differential redshift bins  constrain $\gamma$ more strongly than others.  
The left panel
of Figure~\ref{dvar:fig} shows $\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$.
In both cases the shape of the curve is the same, with the amplitude monotonically decreasing from $z=0.01$ out to $z=0.2$.  If we had
to sculpt the distribution (say due to limited follow-up resources), the preference would be to sculpt out the highest redshift bins
resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the SN-discovery distribution and the sample is unambiguously
parameterized by $z_{max}$.  Note that this result is based on constraining a constant redshift-independent $\gamma$.
\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/dvardz.png}
\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{Left: $\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$.  Right: $\partial \sigma_\gamma / \partial \ln{t}$ as a function of
 $z_{max}$ for two- and ten-year surveys.
\label{dvar:fig}}
\end{figure}
\item[Redshift Depth $z_{max}$]
Increasing the survey redshift depth increases the $\gamma$ precision though that precision asymptotes
$z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
 \item[Survey duration $t$] There is a significant decrease on $\sigma_\gamma$ when going from a 2- to ten-year survey, as shown
 in  Figure~\ref{var:fig}.   The right panel of Figure~\ref{dvar:fig} shows $\partial \sigma_\gamma / \partial \ln{t}$ as a function of
 $z_{max}$ for two- and ten-year surveys.  The non-zero amplitudes show that shot-noise affects both two- and ten-year surveys, but
 much less so for the latter more so at low redshifts.  After 10-years, the effect of sample-variance important for shallower surveys.
\item[Intrinsic Magnitude Dispersion $\sigma_M$]
The relationship between intrinsic magnitude dispersion and precision in $\gamma$ can be seen through the 
previous result, as $\partial \sigma_\gamma/\partial \sigma_M = \partial \sigma_\gamma / \partial \ln{t}$.
Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
Thus, a decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
%limited so the $\sigma^2_M/n$-term
%in the Fisher matrix is important.
Survey follow-up strategy,
which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or equivalently in survey duration. 
 \end{description}

@@

 Shot noise
is important  Figure~\ref{power:fig}.

%
%
%The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
%Nevertheless, their low intrinsic-magnitude uncertainties 
%can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
%of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}. 
%
\section{Peculiar Velocity Surveys are Sensitive to the Growth of Structure $\lowercase{f}D$}
In a peculiar-velocity survey of SN~Ia host galaxies, each galaxy has an observed redshift  and a SN-Ia-based distance.
The distance maps onto a cosmological redshift given a model for the background expansion $H_0$-free.
The  (radial) peculiar velocity is obtained from the difference due to the Doppler effect between the observed and cosmological redshifts.
The statistical properties of the peculiar velocities are summarized through their spatial covariance or  their power spectrum $P_{vv}$.

It is standard practice to express peculiar velocities using the formalism created to describe mass overdensities.
In linear theory, the overdensity field can be expressed as
$
\delta (\mathbf{x},a)  = D(a) \delta (\mathbf{x})$,
where $D(a)$ is  the linear growth factor.
The velocity field is related to the density field through the continuity equation as $v \propto fD\delta$,
where $f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate.  The velocity power spectrum $P_{vv} \propto (fD)^2 P_{\delta \delta}$.
Peculiar velocity surveys, combined with a precision measurement of the power spectrum $P_{\delta \delta}$ 
from the CMB, provide a measure of $fD$
(or  equivalently $f\sigma_8$\footnote{
The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
commonly used in place of $D$ to normalize the
overall amplitude of  overdensities.}).

The survey's   peculiar velocity signal  can be described by its power spectrum $P_{vv}$.  
Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$,
though this detail is not considered here.
The top panel of Figure~\ref{power:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
at $z=0$ expected from General Relativity and linear theory. Ranges of $k$ accessed by surveys with different
maximum redshift $z_{max}$ are also indicated, where $k_{min} \approx (2 z_{max})^{-1}$  and 
$k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$ represents  the smallest
scale where  theory predictions are reliable.
Figure~\ref{power:fig} shows  that the bulk of the peculiar velocity signal comes from $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes.

The probative power of a specific peculiar-velocity tracer primarily depends on its number density and the precision to which its absolute magnitudes are known.
The current generation of peculiar velocity studies uses $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
2016AJ....152...50T}.  These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag.

The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
Nevertheless, their low intrinsic-magnitude uncertainties 
can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.  Existing SN~Ia samples
have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
\citep{2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.

>>>>

In the next decade, SNe~Ia will be used as peculiar-velocity probes to measure  the influence of gravity on structure formation within the Universe.
The peculiar velocity power spectrum is sensitive to the growth of structure as $P_{vv}\propto (fD)^2$, where $D$ is  the linear growth factor and
$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.\footnote{
To be precise, the peculiar
velocity power spectrum also depends on the Hubble parameter as $P_{vv}\propto (HfD)^2$.  A supernova  survey  measures luminosity distance fluctuations
$\delta_{d_L} = (d_L- \bar{d}_L(z))/\bar{d}_L(z)$, where $d_L$ is the observed distance and $\bar{d}_L(z)$ is the expected distance at the observed redshift $z$.
To first order in peculiar velocity along the line of sight $v$, $\delta_{d_L} =  v \left(1-\frac{1}{H\bar{d}_L(z)}\right) \approx -\frac{v}{H\bar{d}_L(z)}$
at low redshift.   The $H$-dependences of  $P_{vv}$ and the conversion from distances to velocities cancel, making  peculiar velocity surveys sensitive to $(fD)^2$.
}
The $\Lambda$CDM prediction for the $z=0$ velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
\citet{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravities 
follow the relation
$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively.  Adopting this parameterization to model gravity, peculiar velocity
surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose $\Lambda$CDM prediction is plotted 
in Figure~2 of  \citet{1475-7516-2013-04-031}.

\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/new2.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{Volume-weighted velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu=1, 0.5$ (solid black, cyan)  as predicted for General Relativity in the linear regime.
Volume-weighted peculiar velocity shot-noise  $k^3\sigma^2/n$ at $z=0.1$, expected from 2- and 10-year (red, black) supernova densities, 0.08 and 0.15~mag (dashed, dotted) intrinsic magnitude dispersions.
The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
redshift depths $z_{max}$.
\label{power:fig}}
\end{figure}

@@

Two advances in the upcoming decade will make SN~Ia peculiar velocities more powerful.
First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\sigma_M \gtrsim 0.12$~mag absolute magnitude
dispersion.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
Recent studies indicate that with the right data, SNe absolute
magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.  Though not yet
established, it is anticipated that such a reduction in intrinsic dispersion comes with a reduction in the magnitude bias correlated with host-galaxy properties
that is observed using current calibrations.
One such SN is worth $\gtrsim 25$ galaxies with 0.4~mag absolute magnitude uncertainty.
Secondly,  in the upcoming decade cadenced wide-field imaging surveys such as ZTF and LSST
  will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia
 for which good light curves can be measured, corresponding to a  number density of $n \sim 5\times 10^{-4}h^3$\,Mpc$^{-3}$.
  This sample has comparable
 number density and more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.

@@

Our scientific interest is in probing gravity so we here focus  on $\gamma$ rather than on $fD$.
To illustrate the distinction,
 $\frac{d(\ln{fD})}{d\gamma} = \ln\Omega + \int \Omega^{\gamma} \ln{\Omega}\, d\ln{a} \approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ respectively
 in $\Lambda$CDM;  two surveys with the same fractional precision  in $fD$ will have different precision in $\gamma$,
 with the one at lower redshift providing the tighter constraint. 
We project uncertainties in $\gamma$, $\sigma_\gamma$ for a suite of idealized surveys
using a Fisher matrix analysis similar to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}.
The Fisher information matrix is
\begin{align}
F_{ij} 
%& = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
%\frac{\partial C}{\partial \lambda_j} \right]\\
& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
\label{fisher:eqn}
\end{align}
where
\begin{equation}
C(k,\mu)  =
  \begin{bmatrix}
   P_{\delta \delta}(k,\mu) + \frac{1}{n} &
   P_{v\delta}(k,\mu)  \\
   P_{v\delta}(k,\mu)  &
  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
   \end{bmatrix}
\label{cov:eq}
\end{equation}
and the two parameters considered are $\lambda \in \{\gamma,bD\}$.
The parameter dependence enters through $(fD)(\gamma)$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy overdensity
power spectrum $P_{\delta \delta }\propto (bD + fD\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
\propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu$ is the cosine of the angle between the $k$-mode
and the line of sight.  While the $bD$ term does contain information on $\gamma$, it is not used for making constraints.
The uncertainty in $\gamma$ is $\sigma_\gamma = \sqrt{\left(F^{-1}\right)_{\gamma\gamma}}$.
Non-GR models may  also predict a change in the scale-dependence of the growth or non-constant $\gamma$, such observations provide additional leverage in probing gravity
but are not considered here.

The uncertainty  $\sigma_\gamma$ of a survey depends on its solid angle $\Omega$, depth $r_{max}$ (or $z_{max}$), duration
$t$ through $n=\epsilon rt$ where $r$ is the observer-frame SN~Ia rate
and $\epsilon$ is the sample-selection efficiency, and
the intrinsic SN~Ia magnitude dispersion 
through $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$.
An analysis in which the RSD and velocity surveys are treated independently is achieved by
setting the off-diagonal elements of $C$ to zero.

We consider peculiar velocity surveys for a range of redshift depths $z_{max}$ for survey durations of $t=2$ and 10~years.
The other survey parameters
$\Omega=3\pi$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
The $k$-limits are  taken to be $k_{min} = (2 r_{max})^{-1}$  and 
$k_{max} = 0.1$~$h$\,Mpc$^{-1}$.
The input bias of SN~Ia host galaxies is set as $b=1.2$.

All the surveys considered can provide meaning tests of gravity.
The projected uncertainty in $\gamma$ achieved by the suite of surveys are shown in Figure~\ref{var:fig}.
For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities, one that uses
both overdensities (RSD) and peculiar velocities independently, and one  that uses
both RSD, peculiar velocities, and their cross-correlation.
The short and shallow,  2-year, $z_{max}=0.1$ survey has $\sigma_\gamma \sim 0.038$,
which can distinguish  between  General Relativity, $f(R)$,  and DGP gravities at the $2\sigma$ level.
The 10-year survey performance asymptotes at $z_{max} \sim0.2$ at a precision of  $\sigma_\gamma \sim 0.01$.
The figure also reveals an interesting interplay between RSD and velocity measurements.  Peculiar velocities alone
can account for much of the probative power of the surveys. RSD alone do not provide significant constraints.
However, considering RSD and velocity cross-correlations decreases $\sigma_\gamma$ by  factors of $\sim 0.8$.
There are important
$k$-modes that are  sample variance limited either in  overdensity and/or peculiar velocity who benefit from the
noise suppression engendered by cross-correlations.

\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/var.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{The projected uncertainty in $\gamma$ achieved by two- (red) and ten-year (black) surveys of varying depth $z_{max}$.
For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed) , one that uses
both overdensities (RSD) and peculiar velocities independently (dotted), and one  that uses
both RSD, peculiar velocities, and their cross-correlation (solid).
\label{var:fig}}
\end{figure}

Survey performance is examined in more detail by considering how $\sigma_\gamma$ changes with respect to the 
survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
within a given survey.  
\begin{description}
\item[Solid Angle $\Omega$]
The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$.  The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole could significantly increase
the sky coverage with a corresponding decrease in the variance of $\gamma$.
\item[Differential Redshift Bin $z$] Certain differential redshift bins  constrain $\gamma$ more strongly than others.  
Figure~\ref{dvar:fig} shows $\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$.
In both cases the shape of the curve is the same, with the amplitude monotonically decreasing from $z=0.01$ out to $z=0.2$.  If we had
to sculpt the distribution (say due to limited follow-up resources), the preference would be to sculpt out the highest redshift bins
resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the SN-discovery distribution and the sample is unambiguously
parameterized by $z_{max}$.  Note that this result is based on constraining a constant redshift-independent $\gamma$.
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{src/dvardz.png}
%\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{$\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$. 
\label{dvar:fig}}
\end{figure}
\item[Redshift Depth $z_{max}$]
Increasing the survey redshift depth increases the $\gamma$ precision though that precision asymptotes
$z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
 \item[Survey duration $t$] There is a significant decrease on $\sigma_\gamma$ when going from a 2- to ten-year survey, as shown
 in  Figure~\ref{var:fig}.   The left panel of Figure~\ref{dvar2:fig} shows $\partial \sigma_\gamma / \partial \ln{t}$ as a function of
 $z_{max}$ for two- and ten-year surveys.  The non-zero amplitudes show that shot-noise affects both two- and ten-year surveys, but
 much less so for the latter more so at low redshifts.  After 10-years, the effect of sample-variance important for shallower surveys.
 \begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
\includegraphics[width=0.47\textwidth]{src/dvardsigM.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{Left: $\partial \sigma_\gamma / \partial \ln{t}$  and Right: $\partial \sigma_\gamma / \partial \sigma_M$ as a function of
 $z_{max}$ for two- and ten-year surveys.
\label{dvar2:fig}}
\end{figure}
\item[Intrinsic Magnitude Dispersion $\sigma_M$]
The relationship between intrinsic magnitude dispersion and precision in $\gamma$ can be seen through the 
previous result, as $\partial \sigma_\gamma/\partial \sigma_M = \partial \sigma_\gamma / \partial \ln{t}$.
Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
Thus, a decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
%limited so the $\sigma^2_M/n$-term
%in the Fisher matrix is important.
Survey follow-up strategy,
which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or equivalently in survey duration. 
 \end{description}

@@

 Shot noise
is important  Figure~\ref{power:fig}.
 (the total rest-frame volumetric SN~Ia rate being $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$.)

%
%
%The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
%Nevertheless, their low intrinsic-magnitude uncertainties 
%can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
%of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}. 
%
%\section{Peculiar Velocity Surveys are Sensitive to the Growth of Structure $\lowercase{f}D$}
%In a peculiar-velocity survey of SN~Ia host galaxies, each galaxy has an observed redshift  and a SN-Ia-based distance.
%The distance maps onto a cosmological redshift given a model for the background expansion $H_0$-free.
%The  (radial) peculiar velocity is obtained from the difference due to the Doppler effect between the observed and cosmological redshifts.
%The statistical properties of the peculiar velocities are summarized through their spatial covariance or  their power spectrum $P_{vv}$.
%
%It is standard practice to express peculiar velocities using the formalism created to describe mass overdensities.
%In linear theory, the overdensity field can be expressed as
%$
%\delta (\mathbf{x},a)  = D(a) \delta (\mathbf{x})$,
%where $D(a)$ is  the linear growth factor.
%The velocity field is related to the density field through the continuity equation as $v \propto fD\delta$,
%where $f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate.  The velocity power spectrum $P_{vv} \propto (fD)^2 P_{\delta \delta}$.
%Peculiar velocity surveys, combined with a precision measurement of the power spectrum $P_{\delta \delta}$ 
%from the CMB, provide a measure of $fD$
%(or  equivalently $f\sigma_8$\footnote{
%The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
%commonly used in place of $D$ to normalize the
%overall amplitude of  overdensities.}).
%
%
%The survey's   peculiar velocity signal  can be described by its power spectrum $P_{vv}$.  
%Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$,
%though this detail is not considered here.
%The top panel of Figure~\ref{power:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
%at $z=0$ expected from General Relativity and linear theory. Ranges of $k$ accessed by surveys with different
%maximum redshift $z_{max}$ are also indicated, where $k_{min} \approx (2 z_{max})^{-1}$  and 
%$k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$ represents  the smallest
%scale where  theory predictions are reliable.
%Figure~\ref{power:fig} shows  that the bulk of the peculiar velocity signal comes from $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes.
%
%The probative power of a specific peculiar-velocity tracer primarily depends on its number density and the precision to which its absolute magnitudes are known.
%The current generation of peculiar velocity studies uses $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
%2016AJ....152...50T}.  These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag. 
%
%
%The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
%Nevertheless, their low intrinsic-magnitude uncertainties 
%can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
%of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.  Existing SN~Ia samples
%have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
%\citep{2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
%However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.
%
